---
title: "Group Comparison Analysis"
author: "ChuqiChen"
date: "2025-12-03"
output: html_document
---
```{r}
library(tidyverse)
library(data.table)
library(ggpubr)
library(scales)
library(plotly)
library(htmlwidgets)
```

```{r}
df_raw <- fread("data/youtube_shorts_tiktok_trends_2025.csv", encoding = "UTF-8") %>% as_tibble()

guess_col <- function(patterns, cols) {
  for(p in patterns){
    idx <- grep(p, cols, ignore.case = TRUE)
    if(length(idx) > 0) return(cols[idx[1]])
  }
  NA_character_
}

cols <- names(df_raw)

col_duration   <- guess_col(c("duration_seconds","duration_sec","duration","video_length"), cols)
col_completion <- guess_col(c("completion_rate","completion","completion_pct"), cols)
col_engagement <- guess_col(c("engagement","interaction","engagement_rate","er"), cols)
col_category   <- guess_col(c("category","content_type","topic","genre"), cols)
col_platform   <- guess_col(c("platform","source","site"), cols)
col_views      <- guess_col(c("views","plays","impressions","view_count"), cols)
col_avgwatch   <- guess_col(c("watch","avg_watch","watch_time"), cols)

df <- tibble(
  duration_sec = as.numeric(df_raw[[col_duration]]),
  completion_rate = as.numeric(df_raw[[col_completion]]),
  category = tolower(as.character(df_raw[[col_category]])),
  platform = if(!is.na(col_platform)) tolower(as.character(df_raw[[col_platform]])) else NA_character_,
  engagement = if(!is.na(col_engagement)) as.numeric(df_raw[[col_engagement]]) else NA_real_,
  views = if(!is.na(col_views)) as.numeric(df_raw[[col_views]]) else NA_real_,
  avg_watch_time_sec = if(!is.na(col_avgwatch)) as.numeric(df_raw[[col_avgwatch]]) else NA_real_
) %>%
  mutate(
    completion_rate = ifelse(completion_rate > 1, completion_rate / 100, completion_rate)
  ) %>%
  filter(!is.na(duration_sec), !is.na(completion_rate))

df <- df %>%
  mutate(
    duration_bin = case_when(
      duration_sec < 15 ~ "<15s",
      duration_sec >= 15 & duration_sec < 30 ~ "15-30s",
      duration_sec >= 30 & duration_sec <= 60 ~ "30-60s",
      duration_sec > 60 ~ ">60s",
      TRUE ~ NA_character_
    ),
    duration_bin = factor(duration_bin, levels = c("<15s","15-30s","30-60s",">60s"))
  )

preferred <- c("lifestyle","sport","music")
present_pref <- intersect(preferred, unique(df$category))
if(length(present_pref) == 0){
  present_pref <- df %>% count(category, sort = TRUE) %>% slice_head(n = 3) %>% pull(category)
}

ggplot(df %>% filter(category %in% present_pref),
       aes(x = category, y = completion_rate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.18, alpha = 0.35, size = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2, color = "red") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Completion Rate by Category", x = "Category", y = "Completion Rate") +
  theme_minimal()
```

Description:

This boxplot with jittered points compares the completion rates across different content categories (specifically "lifestyle" and "music" in the example). Each box represents the distribution of completion rates for a category, with the red dot indicating the mean completion rate.

Interpretation:

It helps identify which categories tend to have higher or more consistent completion rates. For example, "music" may show a higher median completion rate than "lifestyle."

```{r}
ggplot(df %>% filter(!is.na(duration_bin)),
       aes(x = duration_bin, y = completion_rate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.3, size = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2, color = "red") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Completion Rate by Duration Bin", x = "Duration Bin", y = "Completion Rate") +
  theme_minimal()
```

Description:

This boxplot visualizes how completion rates vary across different video duration bins: <15s, 15-30s, 30-60s, and >60s. The red dot marks the mean completion rate for each bin.

Interpretation:

Shows whether shorter videos (e.g., <15s) tend to have higher completion rates compared to longer ones. Often, shorter videos retain viewers better.

```{r}
ggplot(df %>% filter(category %in% present_pref),
       aes(x = duration_bin, y = completion_rate, fill = category)) +
  geom_violin(trim = TRUE, alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2, color = "black",
               position = position_dodge(width = 0.9)) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Completion Rate Distribution by Duration and Category",
       x = "Duration Bin", y = "Completion Rate") +
  theme_minimal()
```

Description:

A violin plot showing the distribution of completion rates across duration bins, separated by category ("lifestyle" and "music"). The black dot represents the mean for each category-bin combination.

Interpretation:

Allows comparison of how completion rates are distributed within each duration bin for different categories. For example, "music" videos may maintain higher completion rates even in longer bins.

```{r}
ggplot(df,
       aes(x = duration_sec, y = completion_rate, color = category)) +
  geom_point(alpha = 0.15, size = 0.7) +
  geom_smooth(aes(color = NULL), method = "loess", span = 0.3, se = TRUE) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Completion Rate vs Duration (LOESS)", x = "Duration (seconds)", y = "Completion Rate") +
  theme_minimal()
```

Description:

A scatterplot with a LOESS smoothing curve showing the relationship between video duration (in seconds) and completion rate. Points are colored by category, but the trend line is global.

Interpretation:

Reveals the overall trend: completion rates generally decline as video duration increases. The LOESS curve helps visualize the non-linear relationship.

```{r}
pred_df <- df %>% select(duration_sec, completion_rate) %>% drop_na()

if(nrow(pred_df) >= 20){
  loess_fit <- loess(completion_rate ~ duration_sec, data = pred_df, span = 0.3)
  grid_x <- seq(min(pred_df$duration_sec), quantile(pred_df$duration_sec, 0.99), length.out = 1200)
  preds <- predict(loess_fit, newdata = data.frame(duration_sec = grid_x))
  pred_tbl <- tibble(duration_sec = grid_x, pred_completion = preds) %>% drop_na()
  peak_row <- pred_tbl %>% slice_max(pred_completion, n = 1)

  ggplot(pred_tbl, aes(x = duration_sec, y = pred_completion)) +
    geom_line(size = 0.8) +
    geom_vline(xintercept = peak_row$duration_sec, linetype = "dashed") +
    annotate("text", x = peak_row$duration_sec, y = max(pred_tbl$pred_completion),
             label = paste0("Peak ≈ ", round(peak_row$duration_sec, 1), "s"),
             hjust = -0.1, vjust = -0.3, size = 3.6) +
    scale_y_continuous(labels = percent_format()) +
    labs(title = "Predicted Optimal Duration (LOESS)",
         x = "Duration (seconds)", y = "Predicted Completion Rate") +
    theme_minimal()
}
```

Description:

A line plot derived from a LOESS model predicting completion rate based on duration. A vertical dashed line marks the duration at which the predicted completion rate peaks.

Interpretation:

Indicates the optimal video duration for maximizing completion rate. In the example, the peak is around 5 seconds, suggesting very short videos are most likely to be completed.

```{r}
ggplot(df %>% filter(category %in% present_pref, !is.na(duration_bin)),
       aes(x = duration_bin, y = completion_rate)) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point", size = 2, color = "red") +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ category) +
  labs(title = "Completion Rate by Duration Bin (Faceted by Category)",
       x = "Duration Bin", y = "Completion Rate") +
  theme_minimal()
```

Description:

A faceted boxplot showing completion rates across duration bins, separately for each category ("lifestyle" and "music"). Red dots indicate bin means.

Interpretation:

Enables category-specific analysis of how completion rates change with duration. For instance, "music" may show less decline in longer bins compared to "lifestyle."