---
title: "EDA for Completion Rate"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
```


```{r}
## delete
shorts_data = read.csv("data/youtube_shorts_tiktok_trends_2025.csv") |> 
  as_tibble()

df = shorts_data |> 
  dplyr::select(
    completion_rate,
    upload_hour,
    is_weekend,
    views,
    likes,
    comments,
    shares,
    saves,
    duration_sec,
    event_season,
    has_emoji,
    creator_tier,
    creator_avg_views,
    platform
  ) |> 
  mutate(
    has_emoji = as.integer(as.character(has_emoji)),
    creator_tier = as.factor(creator_tier),
    platform = as.factor(platform),
    event_season = as.factor(event_season),
    log_views = log(views + 1),
    log_likes = log(likes + 1),
    log_comments = log(comments + 1),
    log_shares = log(shares + 1),
    log_saves = log(saves + 1),
    log_creator_avg = log(creator_avg_views + 1)
    
  ) |> 
  drop_na()
```


### 2.1 Distribution of Outcome (completion rate){#dist-outcome}

Before fitting models, we inspect the distribution of the target variable (*completion_rate*). This helps assess skewness and understand the modeling target's general behavior.

```{r}
# Distribution of outcome variable (completion_rate)
completion_rate_plot <- ggplot(df, aes(x = completion_rate)) +
  geom_histogram(bins = 30, fill = "#9E9AC8", alpha = 0.8, 
                 color = "#54278F", size = 0.4) +
  geom_vline(aes(xintercept = mean(completion_rate)), 
             color = "#E6550D", linetype = "dashed",linewidth = 1) +
  labs(title = "Distribution of Completion Rate",
       x = "Completion Rate", 
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

print(completion_rate_plot)
```

The distribution of completion_rate is roughly centered around 0.6–0.7 with no extreme skewness or heavy tails. Although the shape is not perfectly symmetric, it does not exhibit severe outliers or truncation.



[Next: Regression Analysis →](regression_analysis.html)

