---
title: "EDA for Completion Rate"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 2.1 Distribution of Outcome (completion rate){#dist-outcome}

Before fitting models, we inspect the distribution of the target variable (*completion_rate*). This helps assess skewness and understand the modeling target's general behavior.

```{r}
# Distribution of outcome variable (completion_rate)
completion_rate_plot <- ggplot(df, aes(x = completion_rate)) +
  geom_histogram(bins = 30, fill = "#9E9AC8", alpha = 0.8, 
                 color = "#54278F", size = 0.4) +
  geom_vline(aes(xintercept = mean(completion_rate)), 
             color = "#E6550D", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Completion Rate",
       x = "Completion Rate", 
       y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

print(completion_rate_plot)
```

The distribution of completion_rate is roughly centered around 0.6–0.7 with no extreme skewness or heavy tails. Although the shape is not perfectly symmetric, it does not exhibit severe outliers or truncation.
